---
title: "cnbrrr: Interface to Czech National Bank ARAD Data API"
format: 
  gfm:
    toc: false
author: "Petr Bouchal"
execute:
  eval: true
  echo: true
---

# cnbrrr

R package for accessing data from the Czech National Bank's ARAD (Aggregated data of the CNB) API.

The package is not affiliated in any way to the Czech National Bank but I am grateful to them for proving the data and maintaining the infrastructure.

## About ARAD

[ARAD](https://www.cnb.cz/arad/) is the Czech National Bank's public database containing aggregated economic and financial data. The system provides access to:

- **Monetary policy indicators** - interest rates, money supply, exchange rates
- **Financial market data** - banking sector statistics, capital market indicators  
- **Macroeconomic statistics** - GDP, inflation, employment, government finances
- **Balance of payments** - current account, capital flows, international reserves
- **International comparisons** - harmonized indicators across countries

The data spans from the 1990s to present, with most series updated monthly or quarterly. All data comes from the CNB's statistical processing and external statistical sources.

Comprehensive user guidance is at <https://www.cnb.cz/arad/#/cs/documentation>.

## Package features

The `cnbrrr` package provides:

- **Data discovery**: listing indicators contained in a set or base
- **Metadata exploration** to discover available indicators and categories
- **Simple data retrieval** with automatic date parsing and formatting
- **Consistent data structure** across all functions

## Installation

```r
# Install from GitHub
remotes::install_github("petrbouchal/cnbrrr")
```

## Setup

- obtain an API key from the Czech National Bank (requires registration)
- set it as an environment variable:

```r
Sys.setenv(ARAD_API_KEY = "your_api_key_here")
```

## Quick example

```{r}
library(cnbrrr)

# List available indicators from set (sestava), filter by name
indicators <- arad_list_indicators(set_id = 1032, filter = "domácnost")
indicators
```

Note there is no way to list all available indicators or to list bases and sets - you need to have some idea of what you are looking for - for that, check the website.

```{r}
# Get data for a specific indicator  
data <- arad_get_data("SRUMD08402C")
data
```

See the [Get started](articles/cnbrrr.html) vignette for detailed workflows and examples.

## License

MIT License - see LICENSE file for details.
